---
title: "Text Mining. Comparing corpora"
author: "Aleksei Gorgadze"
output:
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
  code_folding: show
---

## Comparing word usage in contrasting corpora

Texts usually serve as proxy to some
social phenomenon, sentiment, idea, or discourse. In a common research
design, texts from several institutions, groups or actors are
collected. These make two (or more) *contrasting corpora*. Differences
in word usage in these corpora are used to infer something about the
underlying social/political features of the entities these corpora
represent.

But what if we are interested in comparing *content* of the
speech?  How to select words that are interesting? How to avoid
arbitrary choice? To accomplish this, we need some statistical tests
that could show which words are significantly over- or underused in
one corpus **in contrast** to another.

# Install necessary packages, if required. 

```{r, eval=FALSE}
install.packages("readr")
install.packages("dplyr")
install.packages("stringr")
install.packages("tidytext")
install.packages("stopwords")
install.packages("tidyr")
install.packages("ggplot2")

```

## Prepare data

Load State of the Union data:

```{r}
library(readxl)
sou <- read_excel("transcripts_new.xlsx")
warnings()
```

We will now create two corpora for contrasting: speeches by Democratic and
Republican presidents since 1917.

```{r}
library(dplyr)
democrats <- c("Woodrow Wilson", "Franklin D. Roosevelt", "Harry
S. Truman", "John F. Kennedy", "Lyndon B. Johnson", "Jimmy Carter",
"William J. Clinton", "Barack Obama", "Joseph R. Biden, Jr.")
sou.democrats <- sou %>%
    filter(president %in% democrats)
sou.republicans <- sou %>%
    filter(date > "1917-10-25" & !president %in% democrats)
```


## More preprocessing. Stemming and lemmatization

Before we move on to measuring the prevalence of content words, it is
beneficial to make our text data less noisy. In natural language the
same word may take various surface forms: plurals, past tense
etc. This may be of less concern for English, but more so for more
inflected languages. When we are interested in content, it is better
to treat all these *wordforms* as the same word.

There are two common approaches.

1. **Stemming**. Simply remove word endings that contain inflection,
   leaving a *stem*. The approach is simple and fast, and is
   supported by numerous R packages.
2. **Lemmatization**. Reduce a word to its dictionary form — a
   *lemma*. This requires more linguistic data (dictionary etc.), but
   gives more precise results. We will stick to this method.
   

What we need for our task is stems from token-level annotations. This
is how it looks like.

```{r}
library(stringr)
library(tidytext)
library(stopwords)
library(SnowballC)


democrats.long <- sou.democrats %>%
    unnest_tokens(word, transcript) %>%
    mutate(stem = wordStem(word)) %>%
    filter(! word %in% stopwords("en")) %>%
    filter(! str_detect(word, "[0-9]+")) %>%
    mutate(party = "democrat")


republicans.long <- sou.republicans %>%
    unnest_tokens(word, transcript) %>%
    mutate(stem = wordStem(word)) %>%
    filter(! word %in% stopwords("en")) %>%
    filter(! str_detect(word, "[0-9]+")) %>%
    mutate(party = "republican")
```

To make data for comparisons, we now create stem frequency lists for
both corpora and join them in a single table.
Filter out rare words (n<10 for both parties).

```{r}
library(tidyr)
party.stems <- bind_rows(democrats.long, republicans.long) %>%
    dplyr::count(stem, party) %>%
    spread(party, n, fill = 0) %>%
    dplyr::filter(democrat > 10 | republican > 10)
```

## Significance. Dunning log-likelihood (G^2)

The most commonly used statistical measure to evaluate the
significance of the difference of a word's frequency in two corpora is
called log-likelihood (G-squared). 

We start with defining a function that will calculate a list of G2
values given two columns of frequencies (in corpora A and B).

```{r}
g2 = function(a, b) {
  c = sum(a)
  d = sum(b)
  E1 = c * ((a + b) / (c + d))
  E2 = d * ((a + b) / (c + d))
  return(2*((a*log(a/E1+1e-7)) + (b*log(b/E2+1e-7))))
}
```

Now we can calculate log-likelihood for stem frequency
differences for each party:

```{r}
party.g2 <- party.stems %>% 
    mutate(g2=g2(democrat, republican)) %>%
    arrange(desc(g2)) %>%
    mutate(g2 = round(g2, 2))

#View(party.g2)
```

**Question**:

1. Examine words with the most siginificant difference in frequency by
   G2 value. Is this list different from the list of top-frequent
   words? What does it tell us about the word usage in the two subcorpora?
   

Note that the many of the most frequent words turned out to also be
the most significantly disproportional. This should rise suspicions.


## Effect size. Log Ratio

[Friends don't let friends calculate p-values (without understanding them).](http://www.scottbot.net/HIAL/?p=24697)

So we are going to supplement our log-likelihood tests with an effect
size measure that allow to quantify, how large exactly is the difference
of frequencies.

We will use the most conceptually simple effect size measure: Log odds
ratio (See more details and examples in [Text mining with R
book](https://www.tidytextmining.com/twitter.html#comparing-word-usage)). 

Here we define a function similar to g2, and apply it to our data. 

```{r}
logratio <- function(a, b) {
    return(log2((a/sum(a+0.000000000000001)/(b/sum(b+0.000000000000001))+0.000000000000001)))
}
```

Now we may add logrations to our table.

```{r}
party.lr <- party.g2 %>%
    mutate(logratio = logratio(democrat, republican))
```

Words used evenly by both parties:

```{r}
party.lr %>%
    arrange(abs(logratio))
```

Words most acutely overused by one or the other party:

```{r}
party.lr %>%
    filter(democrat > 0 & republican > 0) %>%
    group_by(logratio < 0) %>%
    top_n(15, abs(logratio)) %>%
    ungroup()
```

The same result in a nice looking plot.

```{r}
#install.packages("ggplot2")
library(ggplot2)
party.lr %>%
    filter(democrat > 0 & republican > 0) %>%
    group_by(logratio < 0) %>%
    top_n(15, abs(logratio)) %>%
    ungroup() %>%
    mutate(stem = reorder(stem, logratio)) %>%
    ggplot(aes(stem, logratio, fill = logratio > 0)) +
    geom_col(show.legend = FALSE) +
    coord_flip() +
    ylab("log odds ratio (Republicans/Democrats)") +
    scale_fill_discrete(name = "", labels = c("Democrats", "Republicans"))
```

**Question**:

1. For which of these highly disproportional words we also have enough
evidence to claim statistical significance of the difference?

**Your turn**:

1. Replace stemming with lemmatization using `lemmatize_words`
   function from the `textstem` package.
2. Name the lemmatized column `lem`.
3. Compare lemmatized texts of President Barack Obama and other presidents.
4. Draw a graph and describe the results in a couple of sentences.

```{r}
#replace stemming with lemmatization and name column name
#install.packages("textstem")
library(dplyr)
library(tidytext)
library(stringr)
library(stopwords)
library(tidyr)
library(ggplot2)
library(textstem)

obama.long <- sou %>%
  filter(president == "Barack Obama") %>%
  unnest_tokens(word, transcript) %>%
  mutate(lem = lemmatize_words(word)) %>%
  filter(!lem %in% stopwords("en")) %>%
  filter(!str_detect(lem, "^[0-9]+$")) %>%
  transmute(lem, group = "Obama")

others.long <- sou %>%
  filter(president != "Barack Obama") %>%
  unnest_tokens(word, transcript) %>%
  mutate(lem = lemmatize_words(word)) %>%
  filter(!lem %in% stopwords("en")) %>%
  filter(!str_detect(lem, "^[0-9]+$")) %>%
  transmute(lem, group = "others")
```
```{r}
# frequency lemm
lem.freq <- bind_rows(obama.long, others.long) %>%
  count(lem, group) %>%
  spread(group, n, fill = 0) %>%
  filter(Obama > 10 | others > 10)
```

```{r}
# create function
g2 = function(a, b) {
  c = sum(a)
  d = sum(b)
  E1 = c * ((a + b) / (c + d))
  E2 = d * ((a + b) / (c + d))
  return(2*((a*log(a/E1+1e-7)) + (b*log(b/E2+1e-7))))
}
logratio <- function(a, b) {
    return(log2((a/sum(a+0.000000000000001)/(b/sum(b+0.000000000000001))+0.000000000000001)))
}

lem.stats <- lem.freq %>%
  mutate(
    g2 = g2(Obama, others),
    p  = 1 - pchisq(g2, df = 1),
    p_adj = p.adjust(p, method = "BH"),  
    lr = logratio(Obama, others)
  ) %>%
  arrange(desc(abs(lr)))
```

```{r}
#lemmatized
sig_table <- lem.stats %>%
  filter(p_adj < 0.05) %>%
  top_n(30, wt = abs(lr)) %>%
  arrange(desc(abs(lr))) %>%
  select(lem, Obama, others, lr, g2, p, p_adj)

sig_table 
```
```{r}
# draw a graph
top_lr_plot <- lem.stats %>%
  filter(Obama > 0 & others > 0) %>%
  group_by(lr < 0) %>%
  top_n(15, wt = abs(lr)) %>%
  ungroup() %>%
  mutate(lem = reorder(lem, lr))

ggplot(top_lr_plot, aes(lem, lr, fill = lr > 0)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  xlab("Lemma") +
  ylab("log odds ratio") +
  ggtitle("Обама vs остальные президенты") +
  scale_fill_manual(values = c("FALSE" = "#B0B3B8",  # others
                               "TRUE"  = "#0057B8")) # Obama

```

Сравнение лемматизированных текстов Обамы с речами других президентов показывает системный сдвиг к темам цифровой экономики и климата (lemmas: cyber, tech, online, ceo, entrepreneur, carbon), а также упоминаниям внешнеполитических кейсов его периода (iran’s, afghan). У «остальных» доминируют юридико-институциональные и более архаичные маркеры стиля (shall, treaty, treasury, duty, navy, territory).
